Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-02 paper/tv.tex
m

%as our target language because it Given that Hiphop.js is an integration of Esterel and JavaScript,

Meta-variables are {\textbf{S}}, \code{x} and \code{nm} . Basic signal types include \code{IN} for input signals, \code{OUT} for output signals, \code{INOUT} for both and \code{int} for integer variables. {\textbf{var}} represents the countably infinite set of arbitrary distinct identifiers. We assume that programs are well-typed conforming to basic types \code{\tau}.

A program \code{\mathcal{P} } comprises a list of module definitions \code{ \overrightarrow{module}}. Here, we use the \code{ \overrightarrow{}} script to denote a finite vector (possibly empty) of items.
Each \code{module} has a name \code{nm}, a list of well-typed arguments \code{ \overrightarrow{\tau\ x} } and $ \overrightarrow{\tau\ \textbf{S}}$, a statement-oriented body \code{{p}}, associated with a precondition \code{\effect_{pre}} and a postcondition \code{\effect_{post}}. (The syntax of effects specification \code{\effect} is given in \figref{fig:Syntax_of_Types_and_Effects})

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   {meth^*}} 
\qquad   \qquad
    \m{(Basic\ Types)} \qquad  \code{\tau} \code{  ::= }
     \code{int\ |\ bool\ |\ void}
    \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{\tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ ()\ |\ k^\tau\ |\ x\ |\ 
    %v.f\ |\ 
    %\textbf{new}\ d({v^*}) \ |\ 
    \tau\ x;\ e \ |\ mn({x^*}) \ }
  \code{|\ x{:=}e\ |\ 
 % v.f:=e \ | 
 \ e_1;e_2\ 
  |\ {if}\ v\ {then}\ e_1\ {else}\ e_2 
 }  \\
 &   & & \code{
% |\ e_1}\ op\ \code{ e_2 \ 
| \ \textbf{event} [ev]\
  %|\ \textbf{assert}\ \effect \ 
  | \ \textbf{timeout} \ e \ d \  
  | \ \textbf{deadline} \ e \ d \  
  | \ \textbf{delay}\ d 
  }  \\
  ~\\

@@

@@

@@

\begin{figure}[ht]
    \vspace{-1mm}
    \renewcommand{\arraystretch}{1}
\begin{align*} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          Disjunction                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models \effect_1 {\vee} \effect_2}  
&\m{iff}\quad 
&  d, \varphi \models \effect_1 \ or \ d, \varphi \models \effect_2
\\[0.2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          EMPTY                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models \pi : \empt }  
&\m{iff}\quad 
&  d{=}0 \ and\  {\tt{SAT}(\pi)}  \ and \ \varphi {=} [] 
\\[0.2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          I                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : \anyevent{A} }  &\m{iff}\quad 
& d {\geq} 0 \ and\ {\tt{SAT}(\pi)}  \ and \ 
\varphi {=} \code{[\anyevent{A}]}
  \\[0.2em]
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            CONCATE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : (\es_1{\cdot}\es_2)}  
&\m{iff}\quad 
& {\tt{SAT}(\pi)}  \ and \ \exists \code{\varphi_1, \varphi_2, d_1, d_2}\ and\ \code{
\varphi {=} \varphi_1 {+}{+} \varphi_2, d {=} d_1 {+} d_2 \ and 
}\\
& & &
\code{d_1, \varphi_1 \models  \pi : \es_1}  \ and \ 
\code{d_2, \varphi_2 \models  \pi : \es_2}  
 \\[0.2em]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                       DISJUNCTION                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : (\es_1 {\vee} \es_2)}  
&\m{iff}\quad 
& {\tt{SAT}(\pi)}  \ and \ \code{d, \varphi \models  \pi :  \es_1 } \ or \ 
 \code{d, \varphi \models  \pi :  \es_2}  
\\[0.2em]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                       timed                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : \es \# t }  
&\m{iff}\quad 
& d, \varphi \models  (\pi \wedge t{=}d) : \es
\\[0.2em]
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            KLEENE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d,  \varphi \models \pi : \esn{\es}{\star}  }  
&\m{iff}\quad 
&d, \varphi \models \pi :  \empt \ or \ 
d, \varphi \models \pi :  (\es \cdot \esn{\es}{\star} ) 
 \\[0.2em]
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            FALSE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, \varphi \models False : \bot }  
&\m{iff}\quad 
&   otherwise
%{\tt{UNSAT}(\pi)}  \ or \  \varphi  {=} \bot
%\exists t, d(t) \not\models \llbracket \pure{\pi} \rrbracket_{t}
\end{align*}
    \vspace{-3mm}
\caption{Semantics of \timedEffects.}
\label{fig:Sementic}
  \vspace{-1mm}
\end{figure}

>>>>

Here, we regard \code{k} and \code{x} are meta-variables.  \code{k^\tau} represents a constant of basic type \code{\tau}. \textbf{var} represents the countably infinite set of arbitrary distinct identifiers. 
\anyevent{a} refers to a singleton event coming from the finite set of events \code{\Sigma}. 
%Without loss of generality, any event can be further parametrized with constant values, e.g. \code{\anyevent{a}.0} and  \code{\anyevent{a}}.1 are distinct two events, leading to infinite alphabet.
We assume that programs we use are well-typed conforming to basic types \code{\tau} (we take () as the \code{void} type).
%Each events raising expression \code{\textbf{event}} is parametrized with an event, to trigger such a single event. 
A program \code{\mathcal{P}} comprises a list of 
%data type declarations \code{{datat^*}} and a list of 
method declarations \code{{meth^*}}. 
Here, we use the \code{*} superscript to denote a finite list (possibly empty) of items, for example, \code{{x^*}} refers to a list of variables, \code{x_1, ...,\ x_n}. 
%Besides, \code{d} denotes the name of a user-defined data type, \code{f} denotes a field name.

Each method \code{meth} has a name \code{mn}, an expression-oriented body \code{e}, also is associated with a precondition  \code{\effect_{pre}} and a postcondition \code{\effect_{post}} (the syntax of effects specification \code{\effect} is given in \figref{fig:Syntax_of_Types_and_Effects}).
The language allows each iterative loop to be optimized to an equivalent tail-recursive method, where mutation on parameters is made visible to the caller. The technique of translating away iterative loops is standard
%\citeref{insa2015automatic} 
and is helpful
in further minimising our core language. Expressions comprise unit (), 
constants \code{k}, 
variables \code{x}, 
%one-level field access \code{v.f} (rather than \code{v.f1.f2...} for the sake of simplicity), 
%data structure allocation \code{\textbf{new}\ d({v^*})},  
local variable declaration \code{\tau\ x;\ e}, method calls \code{mn({x^*}) }, 
variable assignments \code{x{:=}e}, 
%field update \code{v.f:=e}, 
expression sequences \code{e_1;e_2},
conditional expressions \code{{if}\ v\ {then}\ e_1\ {else}\ e_2},
event raises expression \code{ \textbf{event} [ev] },

@@

    \m{(Program)} &  \code{\mathcal{P} }&\code{  ::= }&
    \code{%{datat^*}  \  
   {meth^*}} 
\qquad   \qquad\quad 
    \m{(Basic\ Types)} \qquad  \code{\tau} \code{\quad  ::= \quad }
     \code{int\ |\ bool\ |\ void}
    \\

@@

%     \m{(Types)} &  \code{t}&\code{  ::= }&
% \code{d\ |\ \tau} 
% \\    
  \m{(Method\ Def.)} &  \code{meth}&\code{  ::= }&
  \code{\tau \ mn  \ {( \tau\ x)^*}\ \{\textbf{requires}\ \effect_{pre}\ \textbf{ensures}\  \effect_{post}\} \ \{ e \}}
  \\
  \m{(Expressions)} &  \code{e}&\code{  ::= }&
    \code{ ()\ |\ k^\tau\ |\ x\ |\ 
    %v.f\ |\ 
    %\textbf{new}\ d({v^*}) \ |\ 
    \tau\ x;\ e \ |\ mn({x^*}) \ }
  \code{|\ x{:=}e\ |\ 
 % v.f:=e \ | 
 \ e_1;e_2\ 
  |\ {if}\ v\ {then}\ e_1\ {else}\ e_2 
 }  \\
 &   & & \code{
% |\ e_1}\ op\ \code{ e_2 \ 
| \ \textbf{event} [ev]\ 
  | \ \textbf{delay}\ [d] \ 
  | \ e_1\ \textbf{timeout}\  [d] \  e_2 \ 
  | \ \textbf{deadline}  \ [d] \ e \  
  }  \\
  ~\\

@@

A number of timed process constructs can be used to capture common 
real-time system behavior patterns. Without loss of generality, we 
assume d is an integer constant.  $\textbf{delay}\ [d]$ idles for exactly $d$ 
time units. In process $e_1\ \textbf{timeout}\ [d] e_2$, the first observable event of 
$e_1$ shall occur before d time units elapse (since the process starts). 
Otherwise, $e_2$ takes over control after exactly $d$ time units elapse. 
Process $\textbf{deadline}\ [d]\ e$ constrains $e$ to terminate before $d$ time units. 
In this setting, clock variables are 
made implicit and hence they cannot be compared with each other 
directly, which potentially allows efficient clock manipulation and 
hence system verification.

@@

\subsection{Semantics}
\label{subsec:Targetlanguage_Semantics}

In order to define the operational semantics of a system model, 
we define the notion of a configuration to capture the global 
system state during system execution.

\begin{definition}[System configuration]
A system configuration is a pair $c = (V , e)$ where $V$ is a variable valuation function and $e$ is an expression.
\end{definition}

A transition of the system is of the form $c \xrightarrow[]{\text{x}} c^\prime$ where $c$ and $c^\prime$ are 
the system configurations before and after the transition respectively. 
We adopt the following naming convention for transition labels: 
$t$ denotes a non-negative real number; $\tau$ denotes an invisible event;
For instance, $c \xrightarrow[]{\text{t}} c^\prime$ denotes a transition of $t$ time units elapsing. 
In the following, we present the firing rules which are associated with 
the timed process constructs. 
\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Wait                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  t {\leq} d 
}{(V, \textbf{delay}\ [d]) \xrightarrow[]{\text{t}} (V, \textbf{delay}\ [d \text{-}t])}\ [de1]} 
\qquad \qquad 
\code{\frac{
}{(V, \textbf{delay}\ [0]) \xrightarrow[]{\tau} (V, ())}\ [de2]} 
\end{flalign*}

\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   timeout                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (V,e_1) \xrightarrow[]{\text{e}} (V^\prime,e_1^\prime)
}{(V, e_1\ \textbf{timeout} \  [d] \ e_2) \xrightarrow[]{\text{e}} (V^\prime,e^\prime)}\ [to1]} 
\quad  
\code{\frac{
  (V,e) \xrightarrow[]{\tau} (V^\prime,e^\prime)  \quad (t {\leq} d)
}{(V, e_1\ \textbf{timeout} \  [d] \ e_2) \xrightarrow[]{\text{t}} (V, \textbf{timeout}\ e^\prime \ [d \text{-}t])}\ [to2]} 
\\
\code{\frac{
  (V,e_1) \xrightarrow[]{\text{t}} (V,e_1^\prime)  \quad (t {\leq} d) 
}{(V, e_1\ \textbf{timeout} \ [d] \ e_2) \xrightarrow[]{\text{t}} (V,e_1^\prime\ \textbf{timeout} \  [d\text{-}t] \ e_2)}\ [to3]} 
\quad  
\code{\frac{
}{(V, e_1\ \textbf{timeout} \  [0] \ e_2) \xrightarrow[]{\tau} (V, e_2)}\ [to4]} 
\end{flalign*}

\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   deadline                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (V,e) \xrightarrow[]{\text{x}} (V^\prime,e^\prime)
}{(V, \textbf{deadline}\  [d] \ e ) \xrightarrow[]{\text{x}} (V^\prime, \textbf{deadline} \ [d] \ e^\prime)}\ [dl1]} 
\quad  
\code{\frac{
  (V,e) \xrightarrow[]{\text{t}} (V,e^\prime)  \quad (t {\leq} d)
}{(V, \textbf{deadline} \  [d]\ e ) \xrightarrow[]{\text{t}} (V, \textbf{deadline} \ [d \text{-}t] \ e^\prime )}\ [dl2]} 
\end{flalign*}

@@

\begin{figure}[ht]
    \vspace{-1mm}
    \renewcommand{\arraystretch}{1}
\begin{align*} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          Disjunction                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models \effect_1 {\vee} \effect_2}  
&\m{iff}\quad 
&  d, \varphi \models \effect_1 \ or \ d, \varphi \models \effect_2
\\[0.2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          EMPTY                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models \pi : \empt }  
&\m{iff}\quad 
&  d{=}0 \ and\  {\tt{SAT}(\pi)}  \ and \ \varphi {=} [] 
\\[0.2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          I                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : ev }  &\m{iff}\quad 
& d {\geq} 0 \ and\ {\tt{SAT}(\pi)}  \ and \ 
\varphi {=} \code{[ev]}
  \\[0.2em]
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            CONCATE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : (\es_1{\cdot}\es_2)}  
&\m{iff}\quad 
& {\tt{SAT}(\pi)}  \ and \ \exists \code{\varphi_1, \varphi_2, d_1, d_2}\ and\ \code{
\varphi {=} \varphi_1 {+}{+} \varphi_2, d {=} d_1 {+} d_2 \ and 
}\\
& & &
\code{d_1, \varphi_1 \models  \pi : \es_1}  \ and \ 
\code{d_2, \varphi_2 \models  \pi : \es_2}  
 \\[0.2em]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                       DISJUNCTION                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : (\es_1 {\vee} \es_2)}  
&\m{iff}\quad 
& {\tt{SAT}(\pi)}  \ and \ \code{d, \varphi \models  \pi :  \es_1 } \ or \ 
 \code{d, \varphi \models  \pi :  \es_2}  
\\[0.2em]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                       timed                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : \es \# t }  
&\m{iff}\quad 
& d, \varphi \models  (\pi \wedge t{=}d) : \es
\\[0.2em]
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            KLEENE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d,  \varphi \models \pi : \esn{\es}{\star}  }  
&\m{iff}\quad 
&d, \varphi \models \pi :  \empt \ or \ 
d, \varphi \models \pi :  (\es \cdot \esn{\es}{\star} ) 
 \\[0.2em]
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            FALSE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, \varphi \models False : \bot }  
&\m{iff}\quad 
&   otherwise
%{\tt{UNSAT}(\pi)}  \ or \  \varphi  {=} \bot
%\exists t, d(t) \not\models \llbracket \pure{\pi} \rrbracket_{t}
\end{align*}
    \vspace{-3mm}
\caption{Semantics of \timedEffects.}
\label{fig:Sementic}
  \vspace{-1mm}
\end{figure}
